---
title: "ARE 213 Problem Set 2A"
author: "Becky Cardinali, Yuen Ho, Sara Johns, and Jacob Lefler"
date: "Due 10/26/2020"
header-includes: 
  - \usepackage{float}
  - \floatplacement{figure}{H}
  - \usepackage{amsmath}
output: pdf_document
fig_caption: yes
geometry: margin=1.5cm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#======================
# Section 0: Set Up
#======================

# Clear workspace
rm(list = ls())

# Load packages
library(pacman)
library(zoo)
library(haven)
p_load(data.table, dplyr, foreign, readstata13, tidyr, xtable, ggplot2, binom, glmnet, stats, FNN, fastDummies, fixest, parallel)

# set seed for replication
set.seed(1995)
core.num <- (detectCores()/2)
# Directory 
base_directory <- "/Users/sarajohns/Desktop/ARE213_psets/"
# base_directory <- "/Users/beckycardinali/Desktop/ARE213_psets/"
#base_directory <- "/Users/yuen/Documents/GitHub/ARE213_psets/"
# base_directory <- "C:\\Users\\jacob\\Documents\\GitHub\\ARE213_psets\\"

# read data
traffic <- as.data.table(read_dta(paste0(base_directory, "traffic_safety2.dta")))

```

# Question 1

Question 10.3 from Wooldridge: For $T=2$ consider the standard unobserved effects model: 
\begin{equation}
y_{it} = \alpha + x_{it}\beta + c_i + u_{it}
\end{equation}
Let $\hat{\beta}_{FE}$ and $\hat{\beta}_{FD}$ represent the fixed effects and first differences estimators respectively. 

(a) Show that $\hat{\beta}_{FE}$ and $\hat{\beta}_{FD}$ are numerically identical. Hint: it may be easier to write $\hat{\beta}_{FE}$ as the "within estimator" rather than the fixed effects estimator. 

(b) Show that the standard errors of $\hat{\beta}_{FE}$ and $\hat{\beta}_{FD}$ are numerically identical. If you wish, you may assume that $x_{it}$ is a scalar (i.e. there is only one regressor) and ignore any degree of freedome corrections. You are not clustering the standard errors in this problem. 

# Question 2

# Question 3

## A - 

```{r}

# a- Pooled bivariate OLS , yr FE, All covariates 

# create y variable
traffic[, ln_fat_pc := log((fatalities/population))]
# log covariates
traffic[,ln_unemploy := log(unemploy)]
traffic[,ln_totalvmt := log(totalvmt)]
traffic[,ln_precip := log(precip)]
traffic[,ln_snow := log(snow32+0.01)] # to avoid NA from zeroes
# create dummies for FEs (to be used later)
traffic <- dummy_cols(traffic, select_columns = c("year", "state"))

# bivariate OLS
biv <- feols(ln_fat_pc ~ primary + secondary, data=traffic)
summary(biv, se="standard")

biv_yfe <- feols(ln_fat_pc ~ primary + secondary, fixef = "year", data=traffic)
summary(biv_yfe, se = "standard")

biv_yfe_cov <- feols(ln_fat_pc ~ primary + secondary + college + 
                       beer + ln_unemploy + ln_totalvmt + ln_precip +
                       ln_snow + rural_speed + urban_speed, fixef = c("year", "state"), data=traffic)
summary(biv_yfe_cov, se = "standard")

```


## B - 

```{r}

# b - white robust and clustered 

# package command - heteroskedastic
summary(biv, se = "white")
summary(biv_yfe, se = "white")
summary(biv_yfe_cov, se = "white")

# package command - cluster
summary(biv, cluster  = traffic$state)
summary(biv_yfe, cluster = traffic$state)
summary(biv_yfe_cov, cluster = traffic$state)

# write own commands

# will need to get beta matrix manually
calc.beta <- function(xmat, ymat) {
  (solve(t(xmat)%*%xmat)) %*% (t(xmat)%*%ymat)
}

white_middle <- function(xmat, ymat, beta) {
  residsq <- diag(as.vector((ymat - xmat %*% beta)^2))
  mid <- (t(xmat)%*%residsq%*%xmat)
  return(mid)
}

robust.se <- function(xmat, middle) {
  
  var.robust <- solve(t(xmat)%*%xmat) %*% middle %*% solve(t(xmat)%*%xmat)
  
  se <- sqrt(diag(var.robust))
  
  return(se)
}

cluster_middle <- function(i, beta, DT, yvar, xvars) {
  
  state.xmat <- as.matrix(cbind(1,select(DT[state == i,], xvars)))
  state.ymat <- as.matrix(select(DT[state == i,], yvar))
  
  resid <- as.vector(state.ymat - state.xmat %*% beta)
  
  middle.term <- t(state.xmat) %*% resid %*% t(resid) %*% state.xmat
  
  return(middle.term)
}

# List of our variables for the three regressions
biv_var <- c("primary", "secondary")
biv_yfe_var <- c("primary", "secondary", colnames(traffic[,year_1982:year_2003]))
biv_yfe_cov_var <- c("primary", "secondary", "college", "beer",
                     "ln_unemploy", "ln_totalvmt", "ln_precip", 
                     "ln_snow", "rural_speed", "urban_speed", colnames(traffic[,year_1982:year_2003]), 
                     colnames(traffic[,state_2:state_99]))

# Run regression
xmat_biv <- as.matrix(cbind(1,select(traffic, all_of(biv_var))))
xmat_biv_yfe <- as.matrix(cbind(1, select(traffic, all_of(biv_yfe_var))))
xmat_biv_yfe_cov <- as.matrix(cbind(1, select(traffic, all_of(biv_yfe_cov_var))))
ymat <- as.matrix(select(traffic, ln_fat_pc))

beta_biv <- calc.beta(xmat_biv, ymat)
beta_biv_yfe <- calc.beta(xmat_biv_yfe, ymat)
beta_biv_yfe_cov <- calc.beta(xmat_biv_yfe_cov, ymat)

# White robust
# get middle terms
w_mid_biv <- white_middle(xmat_biv, ymat, beta_biv)
w_mid_biv_yfe <- white_middle(xmat_biv_yfe, ymat, beta_biv_yfe)
w_mid_biv_yfe_cov <- white_middle(xmat_biv_yfe_cov, ymat, beta_biv_yfe_cov)
# get standard errors
white_biv <- robust.se(xmat_biv, w_mid_biv)
white_biv_yfe <- robust.se(xmat_biv_yfe, w_mid_biv_yfe)
white_biv_yfe_cov <- robust.se(xmat_biv_yfe_cov, w_mid_biv_yfe_cov)

# Clustered by state
states <- as.vector(unique(traffic[,state]))

cl_mid_biv_terms <- mclapply(states, cluster_middle, beta = beta_biv, DT = traffic, 
                             yvar="ln_fat_pc", xvars=biv_var, mc.cores = core.num)
cl_mid_biv <- Reduce('+', cl_mid_biv_terms)

cl_mid_biv_yfe_terms <- mclapply(states, cluster_middle, beta = beta_biv_yfe, DT = traffic, 
                                 yvar="ln_fat_pc", xvars=biv_yfe_var, mc.cores = core.num)
cl_mid_biv_yfe <- Reduce('+', cl_mid_biv_yfe_terms)

cl_mid_biv_yfe_cov_terms <- mclapply(states, cluster_middle, beta = beta_biv_yfe_cov, DT = traffic, 
                                     yvar="ln_fat_pc", xvars=biv_yfe_cov_var, mc.cores = core.num)
cl_mid_biv_yfe_cov <- Reduce('+', cl_mid_biv_yfe_cov_terms)

cl_biv <- robust.se(xmat_biv, cl_mid_biv)
cl_biv_yfe <- robust.se(xmat_biv_yfe, cl_mid_biv_yfe)
cl_biv_yfe_cov <- robust.se(xmat_biv_yfe_cov, cl_mid_biv_yfe_cov)

```
